---
title: "clustering"
author: "Ari Kamlani"
date: "December 14, 2015"
output: html_document
---

#### Load files and create Tables
```{r, eval=TRUE}
# Load files and Create Tables
TabCarbo  <- read.csv("./AreaWksAllcarbo.txt")
TabCarbof <- TabCarbo[-31, ]
TabEmpty  <- read.csv("./AreaWksAllEmpty.txt")
TabEmptyf <- TabEmpty[-31, ]
TabGlass  <- read.csv("./AreaWksAllglass.txt", header=TRUE, sep="\t", quote="\"", dec=".")
TabGlassf <- TabGlass[-31, ]
TabMelam  <- read.csv("./AreaWksAllmelam.txt", header=TRUE, sep="\t", quote="\"", dec=".")
TabMelamf <- TabMelam[-31, ]
TabWater  <- read.csv("./AreaWksAllWater.txt", header=TRUE, sep="\t", quote="\"", dec=".")
TabWaterf <- TabWater[-31, ]
# create data table
TableAll  <- rbind(TabCarbof, TabEmptyf, TabGlassf, TabMelamf, TabWaterf)
X <- TableAll[, c("RATIO21", "RATIO31")]
colnames(X) <- c("x", "y")
```

#### Clustering
##### Use KMeans for Plotting Centroids and Points for clustering
```{r, eval=TRUE}
# create clusters centers based on mean
C1 <- c(mean(TabCarbof[, c("RATIO21")]), mean(TabCarbof[, c("RATIO31")]))
C2 <- c(mean(TabEmptyf[, c("RATIO21")]), mean(TabEmptyf[, c("RATIO31")]))
C3 <- c(mean(TabGlassf[, c("RATIO21")]), mean(TabGlassf[, c("RATIO31")]))
C4 <- c(mean(TabMelamf[, c("RATIO21")]), mean(TabMelamf[, c("RATIO31")]))
C5 <- c(mean(TabWaterf[, c("RATIO21")]), mean(TabWaterf[, c("RATIO31")]))
center <- t(data.frame(C1,C2,C3,C4,C5))
# perform cluster via kmeans (with outliers), create plots for visualization
km <- kmeans(X, centers=center, iter.max=100, nstart=10)
plot(X, col=km$cluster)
points(km$centers, col = 1:2, pch = 8, cex = 2)
```

##### Plot Clusters via **pam**
```{r, eval=TRUE}
# Alternative Packages: PAM(median,mode), Clara(for >200 observations) Package
# install.packages("fpc")
# install.packages("cluster")
library(fpc)
library(cluster)
set.seed(12152015)
# direct pam usage via input Complete Filtered table
Xprime <- TableAll[, c("RATIO21", "RATIO31")]
colnames(Xprime) <- c("x", "y")
pm <- pamk(Xprime, 5)
cat("Num Clusters: ",pm$nc)
pm$pamobject$medoids
plot(Xprime, col=pm$pamobject$clustering)
points(pm$pamobject$medoids, col = 1:2, pch = 8, cex = 2)
```

##### Plot Clusters via **pam** and **daisy** dissimilarity matrix
```{r, eval=TRUE}
# pam usage based on pairwise dissimilarity matrix
Xdist <- daisy(Xprime, metric="euclidean", stand=FALSE)
Xdist <- as.matrix(Xdist)
pm_dist <- pamk(Xdist, 5)
cat("Num Clusters: ",pm_dist$nc)
# nxn matrix: pm_dist$pamobject$medoids
plot(Xdist, col=pm_dist$pamobject$clustering)
points(pm_dist$pamobject$medoids, col = 1:2, pch = 8, cex = 2)
```

##### Loop usage
- Instead of iterating through each 30 Samples of a clustered table and using a loop to reduce to 10 (30-20) samples for each corresponding table, vectorized operations can be used (e.g. variations of apply on a subsetted dataframe).  Rather than iteratively loop through each of the samples, a vectorized operation on the dataframe could be used.



#### Outliers
- Manually **remove outliers** (keep top 10 as identified by minimized distance (d) from each cluster (Ci) {i:1..k})
```{r, eval=TRUE}
# filter outliers
# manually remove distance points for each centroid and keep 10 best points
filter_outliers <- function(df_in)
{
  for(i in 30:11)
  {
    xhat <- mean(df_in[, c("x")])  
    yhat <- mean(df_in[, c("y")])
    df_in$dx <- abs(xhat - df_in[, c("x")])
    df_in$dy <- abs(yhat - df_in[, c("y")])
    df_in$d  <- sqrt( (df_in$dx)**2 + (df_in$dy)**2)
    df_in    <- df_in[order(df_in$d, decreasing=TRUE), ]
    df_in    <- df_in[-1,]
  }
  return(df_in)
}

exec_outliers <- function(table_in, partition_size=30)
{
  df <- data.frame()
  for (i in seq(1,150, by=30) )
  {
    out <- as.data.frame( filter_outliers(table_in[i:(i+(30-1)),]) )
    out$cluster <- toString(round(i/30)+1)
    df <- rbind(df, out)
  }
  return (df)
}
clustered_filtered_outliers <- exec_outliers(X)
clustered_filtered_outliers
```